A multi-scale story of the diffusion of a new technology: the Web

### 2025-07-29 paper/submission4/anonymised.qmd
submission 3

---
title: "A multi-scale story of the diffusion of a new technology: the web"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol
        department: 
        address: 
        city: 
        state: State
        postal-code: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

@@

<!-- R2 addition -->
The lack of previous granular studies about the diffusion of new technologies 
deprives us from a reliable prior regarding whether the diffusion mechanisms 
and drivers are sensitive or not to the spatial scale that the diffusion of a 
new technology is observed. As per other key concepts in economic geography
[@briant2010dots; @meliciani2015determinants; @andini2019social; @rosenthal2001determinants],
diffusion mechanisms can indeed suffer from Modifiable Areal Unit Problem 
(MAUP): although the adoption of a new technology is at its core an individual 
process, it is hardly ever observed at this, or at a close enough, scale.

@@

<!-- H1 -->
Specifically, two hypotheses are tested. Firstly, whether the three diffusion 
mechanisms namely distance, urban hierarchy and the S-shaped pattern of the 
cumulative level of adoption shaped the diffusion of the commercial Web in 
the UK at granular geographical scales. The analysis firstly illustrates how 
well the S-pattern of cumulative adoption fits at the local scale and offers 
some descriptive insights about the importance of the distance and 
hierarchical diffusion mechanisms. Then, the modelling exercise reveals the 
remarkably good predictive capacity of these mechanisms at granular geographical 
scales. The modelling framework exposes that the relative importance of 
these spatial mechanisms differ across different scales. It also reveals 
spatial heterogeneity. While for some UK regions the importance of the spatial 
diffusion mechanisms are very similar to the rest of the country, other regions
diverge significantly from the country average.

@@

<!-- H2 -->
<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
<!-- and latecomer effects. -->
The second hypothesis assessed whether the diffusion of the web is shaped by 
both early adoption and latecomer effects. On the one hand, the analysis 
highlighted the heterogeneity of adoption rates between LAD as well as the 
spatial clustering of LAD with similar adoption rates. Importantly, the 
observed spatial clustering of LAD with fast diffusion rates can
reinforce existing and rooted patterns of unequal economic productivity 
in the UK [@mccann2022productivity]. Previous research had already indicated 
the long-term productivity gains of the early adoption of such technologies 
[@tranosuk]. On the other hand, the analysis also exposed instances of 
remarkable stability as well as volatility of the Web adoption, with the 
latter being indicative of leapfrogging. While the literature had 
discussed the benefits of technological leapfrogging at the national scale 
[@teece2008firm; @perkins2005international], local scales had escaped such 
studies [@yu2018sustainability]. Thus the leapfrogging evidence presented here 
advocate towards further exploring such effects at granular scales.

The paper offers a number of different contributions. Firstly, it
assesses, for the first time, the role of these spatial diffusion
mechanisms at such granular spatial scales. While previous studies had
highlighted these mechanisms [@hagerstrand1968innovation;
@rogers2010diffusion; @grubler1990rise], their role had never been
tested before at such granular scale. The findings talk directly to this
literature [e.g. @beardsell1999spatial; @haller2011determinants;
@lengyel2020role]. More broadly, the paper offers a robust
methodological framework, which utilises novel data and modern ML
algorithms and techniques to model the diffusion of a new technology
across space and time. It directly addresses the data problem which
prevented economic geographers in engaging with diffusion studies
[@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
seeks to renew economic geography's interest in the diffusion of new
technologies. Current data science advances including the availability
of granular enough data and adequate methods can enable economic
geographers to reveal interesting spatial patterns of diffusion of new
technologies as this paper has demonstrated. This is important as (i)
various stakeholders are interested in predicting localised demand for
new technologies for planning purposes, for example 5G networks
[@leibowicz2016representing;@meade2021modelling], and (ii) the adoption 
of such digital technologies is associated with local economic gains 
[@solow1957technical; @aghion1990model; @kemeny2011international; @tranosuk;
@capello2024nexus]. 
<!-- In addition, while most related studies focused on -->
<!-- tangible technologies, this paper sheds light on the diffusion of a -->
<!-- truly weightless digital technology that is the Web. Thus, it adds much -->
<!-- more granular findings to the limited related studies -->
<!-- [@sinai2004geography; @haller2011determinants].  -->
Importantly, although the geographical focus of this paper is the UK, the 
empirical framework developed here is easily transferable to other geographical 
contexts.

>>>>

---
title: "A multi-scale story of the diffusion of a new technology: the Web"
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

@@

<!-- R2 addition -->
The lack of previous granular studies about the diffusion of new technologies 
deprives us from a reliable prior regarding whether the diffusion mechanisms 
and drivers are sensitive or not to the spatial scale that the diffusion of a 
new technology is observed. As per other key concepts in economic geography
[@briant2010dots; @meliciani2015determinants; @andini2019social],
diffusion mechanisms can indeed suffer from Modifiable Areal Unit Problem 
(MAUP): although the adoption of a new technology is at its core an individual 
process, it is hardly ever observed at this, or at a close enough, scale. 
<!-- response to editor -->
Hence, this paper exposes the importance of spatially-explicit diffusion 
mechanisms at granular geographical scales without the risk of MAUP. The results
are linked to core debates in economic geography about agglomeration 
externalities, centripetal forces and the spatially clustered demand for new 
technologies. Although the paper focuses on the Web, the framework offered here 
and the findings which illustrate the micro-geographical patterns of the 
diffusion of a new technology are applicable to other new technologies.

@@

<!-- H1 -->
Specifically, two hypotheses are tested. Firstly, whether the three diffusion 
mechanisms namely distance, urban hierarchy and the S-shaped pattern of the 
cumulative level of adoption shaped the diffusion of the commercial Web in 
the UK at granular geographical scales. The analysis firstly illustrates how 
well the S-pattern of cumulative adoption fits at the local scale and offers 
some descriptive insights about the importance of the distance and 
hierarchical diffusion mechanisms. Then, the modelling exercise reveals the 
remarkably good predictive capacity of these mechanisms at granular geographical 
scales. It also exposes that the relative importance of these spatial mechanisms 
differ across different scales and the spatial heterogeneity of these findings. 
While for some UK regions the importance of the spatial diffusion mechanisms are 
very similar to the rest of the country, other regions diverge significantly 
from the country average.

@@

<!-- H2 -->
The second hypothesis assessed whether the diffusion of the web is shaped by 
both early adoption and latecomer effects. On the one hand, the analysis 
highlighted the heterogeneity of adoption rates between LAD as well as the 
spatial clustering of LAD with similar adoption rates. 
<!-- response to editor --> 
This empirical finding speaks directly to the spatial clustering literature. 
Just like the clustering of firms [@boschma2011emerging; 
@audretsch1998innovation], the adoption of a new technology is shaped by 
geographically bounded spillovers and path dependency visible even at very
local scales.
Importantly, the observed spatial clustering of LAD with fast diffusion rates 
can reinforce existing and rooted patterns of unequal economic productivity 
in the UK [@mccann2022productivity]. Previous research had already indicated 
the long-term productivity gains of the early adoption of such technologies 
[@tranosuk]. On the other hand, the analysis also exposed instances of 
remarkable stability as well as volatility of the Web adoption, with the 
latter being indicative of leapfrogging. While the literature had 
discussed the benefits of technological leapfrogging at the national scale 
[@teece2008firm; @perkins2005international], local scales had escaped such 
studies [@yu2018sustainability]. Thus the leapfrogging evidence presented here 
advocate towards further exploring such effects at granular scales.

The paper offers a number of different contributions. Firstly, it assesses, for 
the first time, the role of these spatial diffusion mechanisms at granular 
spatial scales. While previous studies had highlighted these mechanisms
[@hagerstrand1968innovation; @rogers2010diffusion; @grubler1990rise], their role 
had never been tested before at such a granular scale 
<!-- response to editor -->
to overcome the potential risk of a MAUP. The spatial heterogeneity of 
the diffusion mechanisms the analysis revealed demonstrates the need for 
observing the diffusion of a new technology at a scale as close as possible to 
the individual level as otherwise significant adoption patterns and, importantly
for economic geography, spatially clustered patterns can be missed. 
The economic geography literature has found evidence of a MAUP in various 
concepts (see for example @briant2010dots for spatial concentration, 
agglomeration economies, and trade determinants and @andini2019social for social 
capital), but this is the first paper that revealed how the scale of observation 
can affect our understanding of the diffusion of a new technology.

In addition, the paper directly addresses the data problem which prevented 
economic geographers in engaging with diffusion studies
[@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
seeks to renew economic geography's interest in the diffusion of new 
technologies.

More broadly, the paper offers a robust methodological framework, which utilises 
novel data and modern ML algorithms and techniques to model the diffusion of a 
new technology across space and time. Current data science advances including 
the availability of granular enough data and adequate methods enable economic
geographers to reveal interesting spatial patterns about the diffusion of new 
technologies as this paper has demonstrated. This is important as (i) various 
stakeholders are interested in predicting localised demand for new technologies 
for planning purposes [@leibowicz2016representing; @meade2021modelling], and 
(ii) the adoption of such digital technologies is associated with local economic 
gains [@solow1957technical; @aghion1990model; @kemeny2011international; 
@tranosuk; @capello2024nexus].

<!-- response to editor -->
These contributions are transferable to different geographical and technological 
contexts. Although the focus of this paper is the Web, the framework and the 
findings are applicable to different technologies. For example, the three 
diffusion mechanisms -- that is distance, urban hierarchy and the S-shaped 
pattern of the cumulative level of adoption -- and their spatial heterogeneity 
can be used to explain and, importantly from a policy perspective, predict local 
demand for different new technologies. While, for instance, it is well expected 
to firstly install new infrastructure such as 5G antennas in central urban areas, 
the results indicate that proximity to these core urban areas can also drive 
local demand for such a technology, probably at a second stage. In terms of 
the empirical framework, it can be used with any point data representing 
technological adoption. Also, archived web data can be obtain for other country 
domains.

<!-- Such an approach could assist  -->
<!-- in mapping the footprints of "technological niches" and, consequently, the role  -->
<!-- of proximity in the creation of such niche markets [@schot2013strategic;  -->
<!-- @bednarz2020pulled].  -->

### 2025-07-29 paper/submission4/anonymised.tex
submission 3

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
  pdfkeywords={diffusion, technologies, Web, digital},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
web}

@@

The lack of previous granular studies about the diffusion of new
technologies deprives us from a reliable prior regarding whether the
diffusion mechanisms and drivers are sensitive or not to the spatial
scale that the diffusion of a new technology is observed. As per other
key concepts in economic geography
\citep{briant2010dots, meliciani2015determinants, andini2019social, rosenthal2001determinants},
diffusion mechanisms can indeed suffer from Modifiable Areal Unit
Problem (MAUP): although the adoption of a new technology is at its core
an individual process, it is hardly ever observed at this, or at a close
enough, scale.

@@

Specifically, two hypotheses are tested. Firstly, whether the three
diffusion mechanisms namely distance, urban hierarchy and the S-shaped
pattern of the cumulative level of adoption shaped the diffusion of the
commercial Web in the UK at granular geographical scales. The analysis
firstly illustrates how well the S-pattern of cumulative adoption fits
at the local scale and offers some descriptive insights about the
importance of the distance and hierarchical diffusion mechanisms. Then,
the modelling exercise reveals the remarkably good predictive capacity
of these mechanisms at granular geographical scales. The modelling
framework exposes that the relative importance of these spatial
mechanisms differ across different scales. It also reveals spatial
heterogeneity. While for some UK regions the importance of the spatial
diffusion mechanisms are very similar to the rest of the country, other
regions diverge significantly from the country average.

@@

The second hypothesis assessed whether the diffusion of the web is
shaped by both early adoption and latecomer effects. On the one hand,
the analysis highlighted the heterogeneity of adoption rates between LAD
as well as the spatial clustering of LAD with similar adoption rates.
Importantly, the observed spatial clustering of LAD with fast diffusion
rates can reinforce existing and rooted patterns of unequal economic
productivity in the UK \citep{mccann2022productivity}. Previous research
had already indicated the long-term productivity gains of the early
adoption of such technologies \citep{tranosuk}. On the other hand, the
analysis also exposed instances of remarkable stability as well as
volatility of the Web adoption, with the latter being indicative of
leapfrogging. While the literature had discussed the benefits of
technological leapfrogging at the national scale
\citep{teece2008firm, perkins2005international}, local scales had
escaped such studies \citep{yu2018sustainability}. Thus the leapfrogging
evidence presented here advocate towards further exploring such effects
at granular scales.

The paper offers a number of different contributions. Firstly, it
assesses, for the first time, the role of these spatial diffusion
mechanisms at such granular spatial scales. While previous studies had
highlighted these mechanisms
\citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
their role had never been tested before at such granular scale. The
findings talk directly to this literature
\citep[e.g.][]{beardsell1999spatial, haller2011determinants, lengyel2020role}.
More broadly, the paper offers a robust methodological framework, which
utilises novel data and modern ML algorithms and techniques to model the
diffusion of a new technology across space and time. It directly
addresses the data problem which prevented economic geographers in
engaging with diffusion studies
\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
paper seeks to renew economic geography's interest in the diffusion of
new technologies. Current data science advances including the
availability of granular enough data and adequate methods can enable
economic geographers to reveal interesting spatial patterns of diffusion
of new technologies as this paper has demonstrated. This is important as
(i) various stakeholders are interested in predicting localised demand
for new technologies for planning purposes, for example 5G networks
\citep{leibowicz2016representing, meade2021modelling}, and (ii) the
adoption of such digital technologies is associated with local economic
gains
\citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.
Importantly, although the geographical focus of this paper is the UK,
the empirical framework developed here is easily transferable to other
geographical contexts.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the Web},
  pdfkeywords={diffusion, technologies, Web, digital},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
Web}

@@

The lack of previous granular studies about the diffusion of new
technologies deprives us from a reliable prior regarding whether the
diffusion mechanisms and drivers are sensitive or not to the spatial
scale that the diffusion of a new technology is observed. As per other
key concepts in economic geography
\citep{briant2010dots, meliciani2015determinants, andini2019social},
diffusion mechanisms can indeed suffer from Modifiable Areal Unit
Problem (MAUP): although the adoption of a new technology is at its core
an individual process, it is hardly ever observed at this, or at a close
enough, scale. Hence, this paper exposes the importance of
spatially-explicit diffusion mechanisms at granular geographical scales
without the risk of MAUP. The results are linked to core debates in
economic geography about agglomeration externalities, centripetal forces
and the spatially clustered demand for new technologies. Although the
paper focuses on the Web, the framework offered here and the findings
which illustrate the micro-geographical patterns of the diffusion of a
new technology are applicable to other new technologies.

@@

Specifically, two hypotheses are tested. Firstly, whether the three
diffusion mechanisms namely distance, urban hierarchy and the S-shaped
pattern of the cumulative level of adoption shaped the diffusion of the
commercial Web in the UK at granular geographical scales. The analysis
firstly illustrates how well the S-pattern of cumulative adoption fits
at the local scale and offers some descriptive insights about the
importance of the distance and hierarchical diffusion mechanisms. Then,
the modelling exercise reveals the remarkably good predictive capacity
of these mechanisms at granular geographical scales. It also exposes
that the relative importance of these spatial mechanisms differ across
different scales and the spatial heterogeneity of these findings. While
for some UK regions the importance of the spatial diffusion mechanisms
are very similar to the rest of the country, other regions diverge
significantly from the country average.

@@

The second hypothesis assessed whether the diffusion of the web is
shaped by both early adoption and latecomer effects. On the one hand,
the analysis highlighted the heterogeneity of adoption rates between LAD
as well as the spatial clustering of LAD with similar adoption rates.
This empirical finding speaks directly to the spatial clustering
literature. Just like the clustering of firms
\citep{boschma2011emerging, audretsch1998innovation}, the adoption of a
new technology is shaped by geographically bounded spillovers and path
dependency visible even at very local scales. Importantly, the observed
spatial clustering of LAD with fast diffusion rates can reinforce
existing and rooted patterns of unequal economic productivity in the UK
\citep{mccann2022productivity}. Previous research had already indicated
the long-term productivity gains of the early adoption of such
technologies \citep{tranosuk}. On the other hand, the analysis also
exposed instances of remarkable stability as well as volatility of the
Web adoption, with the latter being indicative of leapfrogging. While
the literature had discussed the benefits of technological leapfrogging
at the national scale \citep{teece2008firm, perkins2005international},
local scales had escaped such studies \citep{yu2018sustainability}. Thus
the leapfrogging evidence presented here advocate towards further
exploring such effects at granular scales.

The paper offers a number of different contributions. Firstly, it
assesses, for the first time, the role of these spatial diffusion
mechanisms at granular spatial scales. While previous studies had
highlighted these mechanisms
\citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
their role had never been tested before at such a granular scale to
overcome the potential risk of a MAUP. The spatial heterogeneity of the
diffusion mechanisms the analysis revealed demonstrates the need for
observing the diffusion of a new technology at a scale as close as
possible to the individual level as otherwise significant adoption
patterns and, importantly for economic geography, spatially clustered
patterns can be missed. The economic geography literature has found
evidence of a MAUP in various concepts (see for example
\citet{briant2010dots} for spatial concentration, agglomeration
economies, and trade determinants and \citet{andini2019social} for
social capital), but this is the first paper that revealed how the scale
of observation can affect our understanding of the diffusion of a new
technology.

In addition, the paper directly addresses the data problem which
prevented economic geographers in engaging with diffusion studies
\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
paper seeks to renew economic geography's interest in the diffusion of
new technologies.

More broadly, the paper offers a robust methodological framework, which
utilises novel data and modern ML algorithms and techniques to model the
diffusion of a new technology across space and time. Current data
science advances including the availability of granular enough data and
adequate methods enable economic geographers to reveal interesting
spatial patterns about the diffusion of new technologies as this paper
has demonstrated. This is important as (i) various stakeholders are
interested in predicting localised demand for new technologies for
planning purposes \citep{leibowicz2016representing, meade2021modelling},
and (ii) the adoption of such digital technologies is associated with
local economic gains
\citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.

These contributions are transferable to different geographical and
technological contexts. Although the focus of this paper is the Web, the
framework and the findings are applicable to different technologies. For
example, the three diffusion mechanisms -- that is distance, urban
hierarchy and the S-shaped pattern of the cumulative level of adoption
-- and their spatial heterogeneity can be used to explain and,
importantly from a policy perspective, predict local demand for
different new technologies. While, for instance, it is well expected to
firstly install new infrastructure such as 5G antennas in central urban
areas, the results indicate that proximity to these core urban areas can
also drive local demand for such a technology, probably at a second
stage. In terms of the empirical framework, it can be used with any
point data representing technological adoption. Also, archived web data
can be obtain for other country domains.

### 2025-07-29 paper/submission4/title.qmd
submission 3

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Economic Geography}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={diffusion, technologies, Web, digital},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
web}
\author[1]{Emmanouil Tranos%
\corref{cor1}%
}
 \ead{e.tranos@bristol.ac.uk}

\affiliation[1]{organization={University of Bristol, School of
Geographical Sciences},addressline={University Road, BS8
1SS},city={Bristol},country={UK},countrysep={,},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper investigates the spatial diffusion of a new technology that
is the Web in the UK. It employs novel data and machine learning methods
to model the influence of well-established diffusion mechanisms.
Contrary to previous studies, it adopts multiple scales, high spatial
granularity and a long study period that captures the early stages of
the Web until its maturity (1996-2012). Findings reveal the importance
of such spatial mechanisms (namely distance, urban hierarchy and the
S-shaped pattern of the cumulative level of adoption) even at granular
scales. They also highlight spatial heterogeneity and instances of
leapfrogging.
\end{abstract}

\begin{keyword}
    diffusion \sep technologies \sep Web \sep 
    digital
\end{keyword}
\end{frontmatter}

\section{Acknowledgements}\label{acknowledgements}

@@

\section{Funding}\label{funding}

@@

\end{document}

>>>>

---
title: "A multi-scale story of the diffusion of a new technology: the web"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol
        department: School of Geographical Sciences
        address: University Road, BS8 1SS
        city: Bristol
        country: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal of Economic Geography
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

# Acknowledgements

@@

# Funding

@@

### 2025-07-29 paper/submission4/tranos2025.qmd
submission 3

---
title: "A multi-scale story of the diffusion of a new technology: the Web"
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

>>>>

---
title: "A multi-scale story of the diffusion of a new technology: the Web"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol
        department: 
        address: 
        city: 
        state: State
        postal-code: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

### 2025-07-29 paper/submission4/tranos2025.tex
submission 3

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the Web},
  pdfkeywords={diffusion, technologies, Web, digital},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
Web}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the Web},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={diffusion, technologies, Web, digital},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
Web}
\author[1]{Emmanouil Tranos%
\corref{cor1}%
}
 \ead{e.tranos@bristol.ac.uk}

\affiliation[1]{organization={University of
Bristol},postcode={UK},postcodesep={}}
